---
title: "missSBM: Handling missing data in Stochastic Block Models"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[![Travis build status](https://travis-ci.org/jchiquet/missSBM.svg?branch=master)](https://travis-ci.org/jchiquet/missSBM)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/jchiquet/missSBM?branch=master&svg=true)](https://ci.appveyor.com/project/jchiquet/missSBM)
[![Coverage status](https://codecov.io/gh/jchiquet/missSBM/branch/master/graph/badge.svg)](https://codecov.io/github/jchiquet/missSBM?branch=master)

> When a network is partially observed (here, missing dyads, that is, entries with NA in the adjacency matrix rather than 1 or 0), it is possible to account for the underlying process that generates those NAs. *missSBM* is an R package for adjusting the popular Stochastic Block Models from network data  sampled under various missing data conditions.

## Installation

```{r package, eval = FALSE}
devtools::install_github("jchiquet/missSBM")
```

## References

Please cite our work using the following references:

- T. Tabouy, P. Barbillon, J. Chiquet. _Variational Inference for Stochastic Block Models from Sampled Data_, 2018 [link](https://arxiv.org/abs/1707.04141)
