---
title: "missSBM: Handling missing data in Stochastic Block Models"
format:
  gfm:
    default-image-extension: ""
---

<!-- badges: start -->

[![pkgdown](https://github.com/GrossSBM/missSBM/actions/workflows/pkgdown.yaml/badge.svg)](https://grosssbm.github.io/missSBM/) [![R-CMD-check](https://github.com/GrossSBM/missSBM/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/GrossSBM/missSBM/actions/workflows/R-CMD-check.yaml) [![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/missSBM)](https://cran.r-project.org/package=missSBM) [![last commit](https://img.shields.io/github/last-commit/GrossSBM/missSBM.svg)](https://github.com/GrossSBM/missSBM/commits/master) [![Codecov test coverage](https://codecov.io/gh/GrossSBM/missSBM/branch/master/graph/badge.svg)](https://app.codecov.io/gh/GrossSBM/missSBM?branch=master)

<!-- badges: end -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  fig.path  = "man/figures/README-",
  fig.width = 7, fig.height = 5
)
```

## Description

> When a network is partially observed (here, NAs in the adjacency matrix rather than 1 or 0 due to missing information between node pairs), it is possible to account for the underlying process that generates those NAs. 'missSBM', presented in 'Barbillon, Chiquet and Tabouy' (2022) [10.18637/jss.v101.i12](https://doi.org/10.18637/jss.v101.i12), adjusts the popular stochastic block model from network data sampled under various missing data conditions, as described in 'Tabouy, Barbillon and Chiquet' (2019) [10.1080/01621459.2018.1562934](https://doi.org/10.1080/01621459.2018.1562934).

**missSBM** covers a variety of sampling designs for the missing dyads/nodes, split into two families depending on whether the sampling mechanism can be ignored:

- **Missing At Random (MAR)**: `"dyad"`, `"node"`, `"covar-dyad"`, `"covar-node"`, `"snowball"` — whether a dyad/node is observed does not depend on the network itself, so a naive fit on the observed part alone is already consistent.
- **Missing Not At Random (MNAR)**: `"double-standard"`, `"block-node"`, `"block-dyad"`, `"degree"` — the sampling depends on the underlying block structure, the edge value, or the node degree, and must be modeled jointly with the network to avoid biased inference.

## Installation

The last CRAN version is available via

```{r package CRAN, eval = FALSE}
install.packages("missSBM")
```

The development version is available via

```{r package github, eval = FALSE}
devtools::install_github("GrossSBM/missSBM")
```

## Illustration

We illustrate **missSBM** on the French political blogosphere network (`frenchblog2007`, bundled with the package): about 190 blogs, edges standing for hyperlinks, each blog labeled by political party.

### Fit a SBM on the fully observed network

```{r load, warning=FALSE, message=FALSE}
library(missSBM)
library(igraph)

data(frenchblog2007)
frenchblog2007 <- delete_vertices(frenchblog2007, which(degree(frenchblog2007) == 0))
adjacency <- as_adjacency_matrix(frenchblog2007, sparse = FALSE)
```

```{r fit-full, results = 'hide'}
full_fit <- estimateMissSBM(adjacency, vBlocks = 1:10, sampling = "node",
                             control = missSBM_param(trace = FALSE))
```

```{r plot-full}
plot(full_fit, "icl")
```

### Simulate a Missing Not At Random sampling and account for it

Real-world network data collection is rarely complete. We simulate a **block-node** sampling design (MNAR): blogs belonging to smaller political parties are less likely to have been surveyed than blogs in bigger ones, a mechanism tied to the very block structure we are trying to recover -- exactly the kind of design a naive analysis of the observed part alone would get wrong.

```{r observe}
set.seed(1234)
block_prop <- full_fit$bestModel$fittedSBM$blockProp
sampling_rate <- ifelse(block_prop < 0.1, 0.2, 0.8)
partial_network <- observeNetwork(adjacency, sampling = "block-node",
                                   parameters = sampling_rate,
                                   clusters = full_fit$bestModel$fittedSBM$memberships)
mean(is.na(partial_network)) # fraction of missing dyads
```

```{r fit-partial, results = 'hide'}
partial_fit <- estimateMissSBM(partial_network, vBlocks = 1:10, sampling = "block-node",
                                control = missSBM_param(trace = FALSE, iterates = 2))
```

```{r plot-partial}
plot(partial_fit, "icl")
```

Despite about 8% of the dyads missing under this informative design, **missSBM** recovers a block structure close to the one found on the fully observed network (Adjusted Rand Index, the closer to 1 the better):

```{r ari}
aricode::ARI(full_fit$bestModel$fittedSBM$memberships, partial_fit$bestModel$fittedSBM$memberships)
```

```{r plot-expected, fig.width = 10, warning=FALSE}
par(mfrow = c(1, 2))
plot(full_fit$bestModel, "expected", dimLabels = list(row = "blogs", col = "blogs"))
plot(partial_fit$bestModel, "expected", dimLabels = list(row = "blogs", col = "blogs"))
```

See `vignette("case_study_war_networks")` and [the other articles](https://grosssbm.github.io/missSBM/articles/) for a more comprehensive tour, including covariates and further sampling designs.

## References

Please cite our work using the following references:

Barbillon, P., Chiquet, J., & Tabouy, T. (2022). missSBM: An R Package for Handling Missing Values in the Stochastic Block Model. _Journal of Statistical Software_, 101(12), 1–32. DOI: [10.18637/jss.v101.i12](https://doi.org/10.18637/jss.v101.i12)

Timothée Tabouy, Pierre Barbillon & Julien Chiquet (2019) "Variational Inference for Stochastic Block Models from Sampled Data", _Journal of the American Statistical Association_, DOI: [10.1080/01621459.2018.1562934](https://doi.org/10.1080/01621459.2018.1562934)
