---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# missSBM: Handling missing data in Stochastic Block Models

[![Travis build status](https://travis-ci.org/jchiquet/missSBM.svg?branch=master)](https://travis-ci.org/jchiquet/missSBM)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/jchiquet/missSBM?branch=master&svg=true)](https://ci.appveyor.com/project/jchiquet/missSBM)
[![Coverage status](https://codecov.io/gh/jchiquet/missSBM/branch/master/graph/badge.svg)](https://codecov.io/github/jchiquet/missSBM?branch=master)

> When a network is partially observed (here, missing dyads, that is, entries with NAs in the adjacency matrix rather than 1 or 0), it is possible to account for the underlying process that generates those NAs. missSBM adjusts the popular Stochastic Block Model from network data sampled under various missing data conditions, as described in [10.1080/01621459.2018.1562934](https://doi.org/10.1080/01621459.2018.1562934).

## Installation

```{r package, eval = FALSE}
devtools::install_github("jchiquet/missSBM")
```

## Reference

Please cite our work using the following reference:

Timothée Tabouy, Pierre Barbillon & Julien Chiquet (2019) "Variational Inference for Stochastic Block Models from Sampled Data", _Journal of the American Statistical Association_, DOI: [10.1080/01621459.2018.1562934](https://doi.org/10.1080/01621459.2018.1562934)

